Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset

### 2025-06-17 code/manuscript.Rmd
Added graphical abstract and updated the manuscript

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on pre- and post- retrofit comparison to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events which are static changes or more gradual changes denoted as operational drift occur during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies conventional M&V suggested by industry guidelines, the results can be largely confounded. In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method which randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. Our results show that compared to the conventional method, the randomized method provides faster and more accurate savings estimations. Additionally we found that when non-routine events are present, the randomized M&V approach estimates savings that are much closer to the ground-truth savings than the conventional M&V method, demonstrating much improved reliability. We also assessed the impact of starting the M&V at different time of the year, continuing randomization with different sampling ratio after satisfying all criteria, dropping samples affected by carryover when switching between strategies, and identify the optimal sampling interval for that scenario using this large dataset.

@@

@@

To demonstrate and differences between the two M&V methods, the Building
Genome Dataset 2 (BGD2) is used, which is an extensive open-access
dataset designed to advance research and development in building energy
efficiency and control strategies acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 1000 buildings' metadata and realistic operational
measurements between 2016 and 2017 from across North America and Europe,
making it one of the most comprehensive collections of building-related
data available for scientific use. The dataset includes various
commercial building types such as offices, education facilities, public,
and retail buildings, and provides detailed information on their
physical characteristics (e.g. energy ratings, heating types and floor
area) and hourly measurements of chilled and hot water, electricity, gas
usage as well as site outdoor weather conditions. In this study, we
mostly used the metered electricity and outdoor weather measurements for
running the M&V methods.

@@

ASHRAE Guideline 14 and the IPMVP provide options for whole-building M&V
use cases. For code compliance using the prescriptive option, 12 months
of baseline and 12 months of post-retrofit measurements are required. In
addition, the expected savings should be larger than 10% and the
baseline energy model fitting accuracy should have CV(RMSE) lower than
25% for 12-month data collection. Thus, most research related to M&V for
whole-building approach focuses on the accuracy of baseline modeling,
exploring model performance from simple regression models to more
complex machine learning techniques. One study reviewed various models
suitable for M&V applications as well as selected input features
[@alrobaie_review_2022] and another study provided a definitive
methodology to apply machine learning models for M&V use cases
[@gallagher_development_2018]. In addition, a few studies investigated
the critical performance metrics to evaluate the developed baseline
models [@granderson_automated_2015; @granderson_development_2014] and
compared a variety of models using those metrics
[@granderson_accuracy_2016]. These studies made significant
contributions by emphasizing the uncertainty associated with the
model-fitting process, a key factor in accurately determining energy
savings. Furthermore, other researchers addressed this issue by
leveraging statistical formulation and inference to improve baseline
energy models [@burkhart_measurement_2014; @heo_calibration_2012;
@walter_uncertainty_2014]. However, a gap still remains in the
literature regarding the rigorous quantification of uncertainties
directly associated with calculated savings, for instance, accounting
for the potential bias that the baseline model might deteriorate (i.e.
become 'stale') over the extended period of pre- and post-analysis.

@@

Changes in a building that influences its energy usage are commonly
refered to as non-routine events in M&V and are typically unrelated to
the intervention strategy. Current guidelines define non-routine events
as 'static factors' that needed to be adjusted after projecting baseline
in the post-retrofit period [@ashrae_ashrae_2023;
@efficiency_valuation_organisation_international_2012]. A common
non-routine event in energy-saving M&V projects is a change in occupancy
or a significant shift in occupant behavior, equipment run time, and
operating conditions (e.g. set points, lighting, and ventilation
levels). However, current standards or guidelines only provide a general
approach for consideration. For example, IPMVP requires the facility
owner and the M&V analyst to periodically perform inspections of all
equipment and operations during the reporting period, which is
labor-intensive and error-prone. ASHRAE guideline 14 recommends
performing engineering calculations or computer software simulations to
adjust the post-retrofit baseline. Additionally, it is relatively rare
to have access to all measurements needed for adjustments such as
occupancy, and thus the analyst normally assumes those factors remain
unchanged throughout the study. For example, some studies uses linear
interpolation to estimate counter-factual baseline for demand response
program [@beil_round-trip_2015; @keskar_commercial_2020]. Consequently,
one study points out that it is inaccurate to assume no change in the
operating conditions during the response period
[@huang_experimental_2023]. Other studies realized such limitation in
field study, and emphasized the importance of requiring matched groups
to control for exogenous factors beyond weather differences when
comparing between baseline and intervention
[@demand_side_analytics_population_2022; @huang_experimental_2023].

Overall, those non-routine events mostly refer to clearly observable
changes, but there are also more gradual and subtle changes in
buildings, which are hardly noticed. We define those gradual changes as
'operational drift' in this study. One likely encountered operational
drift situation is the filter clogging in air handling units due to
particle accumulation. This causes supply fans to gradually consume more
energy to maintain the required duct static pressure [@feng_newly_2019;
@zhai_full-scale_2017]. If M&V analysts are unaware of such changes and
lack an appropriate adjustment method (e.g., replacing filters before
the intervention begins), the savings could be overestimated as the
decreased energy use is incorrectly attributed to the intervention
rather than the filter replacement. An operational drift has not been
formally defined in the literature or standards, but in this study, we
consider it as a special type of non-routine events as it also
represents a change in facility operation but in a longer term.

@@

As mentioned, the goal of an M&V project is to determine the effect,
typically energy savings, of some intervention in the building. In this
study, we focus on switchable interventions and an example of such an
intervention is a control strategy that adjusts the chilled water
plant's supply water temperature, which can be commonly found in the
literature [@duarte_field_2023; @lee_simulationoptimization_2012;
@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. In our case, we
made it even simpler by adjusting the setpoint based on outdoor weather
conditions with more description shown in Section
\@ref(apply-virtue-intervention). By conducting such an analysis, we aim
to:

@@

```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="75%"}

@@

For all qualified buildings, we further added one more stringent
criterion to explore all buildings with very stable electricity usage
between the two years:

1.  No statistically significant difference (P-value \> 0.05) between
    the two-year electricity usage, which are sometimes assumed to be
    the case for whole-building measurement and verification.

As a result, only 66 buildings are labeled in this 'stable' subset (out
of total of 600 buildings) implying the assumption that buildings are
typically stable over the timeframe involved in conventional M&V is
rarely true. As mentioned in Section \@ref(non-routine-events-impact),
such variability is largely associated with non-routine events,
especially operational drifts since the assessment, even under ideal
situations, last for two years. And according to the statistics shown
here, it is a typical case observed in real buildings applications.

@@

|                        |                                                                  |       |
|-------------------|----------------------------------|-------------------|
| Parameter              | Description                                                      | Value |
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

-   A minimum and maximum of 12 and 108 weeks respectively. The
    randomized schedule covers the entire two-year period but stopping
    criteria enables an early stop at the end of the satisfied blocking
    period.

-   At least 80% of the dry-bulb temperature range in the annual TMY
    data sampled by both strategies.

-   90% confidence that energy savings exceed or do not exceed 0% using
    the SPRT test. Medium effect size (d = 0.5) quantified by cohen's d
    and calculated SPRT statistics either falls below the lower
    threshold or exceeds the upper threshold.

@@

@@

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_hline(yintercept = 0, color = "#fb8072", linewidth = 1.5, lty = "dashed") +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

ggarrange(p_accuracy, p_timeline, 
          ncol = 1, nrow = 2,
          heights = c(2.2, 1),
          labels = c("", "c)"),
          common.legend = F, 
          legend = "bottom") +
  plot_annotation(title = "Overall accuracy and time required for savings estimation by different M&V methods")
```

Figure \@ref(fig:abstract) shows the overall results of M&V methods
comparison and for clarity, we plotted conventional M&V results in grey,
randomized M&V results after satisfying all stopping criteria in green,
and the randomized M&V results sampled for the entire 24 months in blue.
Subplots (a) and (b) calculate the savings estimation error as the
absolute deviation from the true savings (i.e. $|Acc|$) and show the
results distribution using the conventional M&V method in the first
column, the randomized M&V method that stops after satisfying all
stopping criteria in the second column, and the randomized M&V method
continues 50%/50% sampling throughout 2 years in the third column. In
addition, subplot (a) shows the savings estimated from measured weather
conditions, and subplot b) shows the savings normalized on the typical
meteorological weather of the building location after using TOWT model.
As a result, the conventional M&V method exhibits a median deviation of
5% in savings estimation, whereas the proposed randomized method
demonstrates significantly smaller deviations. If the analyst stops
immediately after satisfying all stopping criteria, the deviation is
reduced to approximately 1.6%. Extending the M&V period to match the
length of the conventional method (24 months) further improves accuracy
to 0.7% deviation. Furthermore, by comparing subplots (a) and (b), we
observe a minor improvement in estimation accuracy when a model is
fitted to the data to account for weather differences. The model fitting
is most beneficial for the shorter dataset (randomized and stopping when
all criteria are met, typically 24-36 weeks of data), where it improves
estimation accuracy by 0.4%. This is because the weather conditions
between the first and second year for each of the locations are
relatively similar on average over this time period, and thus the effect
of adjusting for temperature as an independent variable in the model is
relatively small. Subplot (c) compares the overall timeline of the two
M&V methods. The results show that almost all buildings achieve accurate
M&V results and meet the stopping criteria within 24 weeks (for this
intervention), with the remainder meeting all stopping criteria within
36 weeks. The additional 12 weeks is usually driven by a need to span
sufficient weather conditions representative of a full year of data.
Meanwhile, the red arrow highlights the excess time required by the
conventional M&V method (24 months). Thus, the randomized M&V method
finds more accurate results in approximately a quarter of the time
required for conventional M&V. Even if it was the case that an existing
12-month baseline data was available at the start of the M&V process,
randomized M&V would still obtain a more accurate result, in less time,
than obtaining the 12-month retrofit dataset required by conventional
M&V.

We also plotted the estimated deviation without absolute calculation in
Figure \@ref(fig:meanerr) and we noticed that the uncertainty range
associated with the conventional method is significantly larger compared
to the randomized method that only takes 36 weeks according to subplot
(c) in Figure \@ref(fig:abstract).

```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods", fig.height=8, fig.width=12}
# plot for the stable subset
rand_eob_S <- df_seq_fs_stable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

@@

|                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
|-------------------|----------------------------------|-------------------|
| 25th percentile | -5.85                                               | `r percentiles$Q1`     |
| 50th percentile | -1.25                                               | `r percentiles$Median` |
| 75th percentile | 3.86                                                | `r percentiles$Q3`     |

: Table 1. Savings accuracy comparison with a similar literature

@@

@@

As mentioned in Section \@ref(filter-and-clean-dataset), the majority of
the selected buildings are affected by operational drift, which is the
reason why we observe baseline electricity usage changes throughout the
two year period. To evaluate the robustness of the two M&V methods to
the effect of operational drift and non-routine events, we repeated the
M&V run on the original dataset prior to the application of the reset
control. In other words, there is no intervention applied here, so the
most reliable M&V method will be the one that detects closer to 0 kW
savings using the measurements. Therefore, the $Acc$ here are calculated
as:

Figure \@ref(fig:nosaving) shows the randomized M&V method (both at the
time of satisfying all stopping criteria and at the end of a two-year
period) produced consistent results. We included the absolute deviation
($|ACC|$) figure in the supplementary material. Similarly to Figure
\@ref(fig:abstract), stopping when meeting all stopping criteria
generally yields slightly larger uncertainty than continuing for 24
months. However, it still yields a more accurate result than the
conventional method. Interestingly, similar to
[@granderson_accuracy_2016] where they obtained a median normalized
deviation around -1.25%, our analysis on conventional method shows that
the typical deviation is -1.5%, meaning that the typical building is
decreasing energy use slightly in this dataset in the absence of any
known intervention. Thus, if a building analyst were to apply an
intervention to this building and use conventional M&V to measure the
effect, they would 'detect' 1.5% energy savings even when the
intervention had no actual impact on energy performance.

@@

Sampling at different intervals and excluding non-consecutive days also
impact other stopping criteria, such as the time needed to satisfy
statistical uncertainty and weather conditions thresholds. In Figure
\@ref(fig:intervaltime), we compared the percentage of buildings
satisfying all stopping criteria under different sampling intervals,
with or without dropping non-consecutive days. The results indicate that
randomized M&V concludes more quickly when all data are retained,
particularly at shorter intervals. For instance, when sampling at a
one-day interval, 89% of buildings meet all criteria within 24 weeks if
no data are dropped, but this number decreases significantly to 52% when
non-consecutive days (account for 1/3 of the data) are excluded, and 3%
of the buildings require another 12 weeks to satisfy the weather
criterion. As the sampling interval increases, this difference gradually
diminishes: when sampling at a weekly interval, removing non-consecutive
days has no impact on the timeline of meeting stopping criteria.

>>>>

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on comparing pre- and post-retrofit performance to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events (step changes or more gradual changes in how a building operates) occur during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies the conventional M&V method, the results can be confounded. In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method that randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset (639 buildings) covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. The results show that compared to the conventional method, the randomized method provides more accurate savings estimations with a meidum of 74% accuracy improvement and is even faster. Additionally, we found that when non-routine events are present, the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability. We also assessed the impact of normalizing for different weather, starting the M&V at different time of the year, continuing randomization with a different sampling ratio after satisfying all criteria, dropping samples affected by carryover when switching between strategies, and identified the optimal sampling interval for that scenario using this large dataset.

@@

### Graphical abstract {.unnumbered}

```{r graphicalabstract, out.width = "80%"}

knitr::include_graphics("../figs/manuscript/graphical_abs.jpg", dpi = 900)
```

### Keywords {.unnumbered}

Measurement and verification (M&V); Controls; Randomized switchback
design; Uncertainty quantification; Building energy analysis; Heating
ventilation and air conditioning

### Highlights {.unnumbered}

-   Verified the improvements of the randomized switchback method using
    a large open-source dataset
-   Addressed the influence of non-routine events and also more subtle
    long-term 'operational drift' in buildings
-   Investigated the trade-off of selecting longer sampling intervals to
    account for carryover effect and weather dependency of the method
-   Discussed the impact of selecting different sampling ratio on the
    M&V accuracy

@@

To demonstrate and differences between the two M&V methods, the Building
Genome Dataset 2 (BGD2) is used, which is an extensive open-access
dataset designed to advance research and development in building energy
efficiency and control strategies, acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 1000 buildings' metadata and realistic operational
measurements between 2016 and 2017 from across North America and Europe,
making it one of the most comprehensive collections of building-related
data available for scientific use. The dataset includes various
commercial building types such as offices, education facilities, public,
and retail buildings, and provides detailed information on their
physical characteristics (e.g. energy ratings, heating types and floor
area) and hourly measurements of chilled and hot water, electricity, gas
usage as well as site outdoor weather conditions. In this study, we
mostly used the metered electricity and outdoor weather measurements for
evaluating the M&V methods.

@@

ASHRAE Guideline 14 and the IPMVP provide options for whole-building M&V
use cases. For code compliance using the prescriptive option, 12 months
of baseline and 12 months of post-retrofit measurements are required. In
addition, the expected savings should be larger than 10% and the
baseline energy model fitting accuracy should have CV(RMSE) lower than
25% for 12-month data collection. Thus, most research related to M&V for
the whole-building approach focuses on the accuracy of baseline
modeling, exploring model performance from simple regression models to
more complex machine learning techniques. One study reviewed various
models suitable for M&V applications as well as selected input features
[@alrobaie_review_2022] and another study provided a definitive
methodology to apply machine learning models for M&V use cases
[@gallagher_development_2018]. In addition, a few studies investigated
the critical performance metrics to evaluate the developed baseline
models [@granderson_automated_2015; @granderson_development_2014] and
compared a variety of models using those metrics
[@granderson_accuracy_2016]. These studies made significant
contributions by emphasizing the uncertainty associated with the
model-fitting process, a key factor in accurately determining energy
savings. Furthermore, other researchers addressed this issue by
leveraging statistical formulation and inference to improve baseline
energy models [@burkhart_measurement_2014; @heo_calibration_2012;
@walter_uncertainty_2014]. However, a gap still remains in the
literature regarding the rigorous quantification of uncertainties
directly associated with calculated savings, for instance, accounting
for the potential bias that the baseline model might deteriorate (i.e.,
become 'stale') over the extended period of pre- and post-analysis.

@@

Changes in a building that influence its energy usage and are unrelated
to the intervention being studied by the M&V method are commonly
referred to as non-routine events. Current guidelines define non-routine
events as 'static factors' that need to be adjusted after projecting the
baseline in the post-retrofit period [@ashrae_ashrae_2023;
@efficiency_valuation_organisation_international_2012]. A common
non-routine event in energy-saving M&V projects is a change in occupancy
or a significant shift in occupant behavior, equipment run time, and
operating conditions (e.g. set points, lighting, and ventilation
levels). However, current standards or guidelines only provide a general
approach for consideration. For example, IPMVP requires the facility
owner and the M&V analyst to periodically perform inspections of all
equipment and operations during the reporting period, which is
labor-intensive and error-prone. ASHRAE Guideline 14 recommends
performing engineering calculations or computer software simulations to
adjust the post-retrofit baseline. Additionally, it is relatively rare
to have access to all measurements needed for adjustments, such as
occupancy, and thus the analyst normally assumes those factors remain
unchanged throughout the study. For example, some studies use linear
interpolation to estimate counter-factual baseline for the demand
response program [@beil_round-trip_2015; @keskar_commercial_2020].
Consequently, one study points out that it is inaccurate to assume no
change in the operating conditions during the response period (Huang,
Katipamula, and Lutes 2023). Other studies realized such limitation in
field study, and emphasized the importance of requiring matched groups
to control for exogenous factors beyond weather differences when
comparing between baseline and intervention
[@demand_side_analytics_population_2022; @huang_experimental_2023].

Overall, those non-routine events mostly refer to clearly observable
changes, but there are also more gradual and subtle changes in
buildings, which are hardly noticed. We define those gradual changes as
'operational drift' in this study. One likely encountered operational
drift situation is the filter clogging in air handling units due to
particle accumulation. This causes supply fans to gradually consume more
energy to maintain the required duct static pressure [@feng_newly_2019;
@zhai_full-scale_2017]. If M&V analysts are unaware of such changes and
lack an appropriate adjustment method (e.g., replacing filters before
the intervention begins), the savings could be overestimated as the
decreased energy use is incorrectly attributed to the intervention
rather than the filter replacement. An operational drift has not been
formally defined in the literature or standards, but in this study, we
consider it as a special type of non-routine event, as it also
represents a change in facility operation that occurs gradually over a
longer period of time.

@@

As mentioned, the goal of an M&V project is to determine the effect,
typically energy savings, of some intervention in the building. In this
study, we focus on switchable interventions and an example of such an
intervention is a control strategy that adjusts the chilled water
plant's supply water temperature, which can be commonly found in the
literature [@duarte_field_2023; @lee_simulationoptimization_2012;
@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. In our case, we
made it even simpler by adjusting the setpoint based on outdoor weather
conditions (see description in Section
\@ref(apply-virtue-intervention)). By conducting such an analysis, we
aim to:

@@

```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="70%"}

@@

For all qualified buildings, we also create a feature to identify
buildings with very stable electricity usage overall by identifying
those with no statistically significant difference (P-value \> 0.05)
between the two-year electricity usage. Applying this filter to real
buildings, only 66 buildings are labeled in this 'stable' subset out of
a total of more than 600 buildings. While we did not adjust for weather
differences between the two years when generating this high-level
feature, this result highlights that the assumption that buildings are
typically stable over the timeframe involved in conventional M&V methods
is rarely true. As mentioned in \@ref(non-routine-events-impact), such
variability is largely associated with non-routine events, especially
operational drifts since the assessment, even under ideal situations,
lasts for two years. And according to the statistics shown here, it is a
typical case observed in real buildings applications.

@@

| Parameter              | Description                                                      | Value |
|------------------|------------------------------------|------------------|
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

-   A minimum and maximum of 12 and 108 weeks respectively. The
    randomized schedule covers the entire two-year period but the
    stopping criteria enable an early stop at the end of the satisfied
    blocking period.

-   At least 80% of the dry-bulb temperature range in the annual TMY
    data is sampled by both strategies.

-   90% confidence that energy savings exceed or do not exceed 0% using
    the sequential probability ratio test (SPRT). Medium effect size (d
    = 0.5) quantified by Cohen's d and calculated SPRT statistics either
    fall below the lower threshold or exceed the upper threshold.

@@

Figure \@ref(fig:abstract) shows the overall results of M&V methods
comparison and for clarity, we plotted conventional M&V results in grey,
randomized M&V results after satisfying all stopping criteria in green,
and the randomized M&V results sampled for the entire 24 months in blue.
Subplots (a) and (b) calculate the savings estimation error as the
absolute deviation from the true savings (i.e. $|Acc|$) and show the
results distribution using the conventional M&V method in the first
column, the randomized M&V method that stops after satisfying all
stopping criteria in the second column, and the randomized M&V method
continues 50%/50% sampling throughout 2 years in the third column. In
addition, subplot (a) shows the savings estimated from measured weather
conditions, and subplot b) shows the savings normalized on the typical
meteorological weather of the building location after using a
time-of-week-temperature (TOWT) model [@mathieu_quantifying_2011]. As a
result, the conventional M&V method exhibits a median deviation of 5% in
savings estimation, whereas the randomized method demonstrates
significantly smaller deviations. If the analyst stops immediately after
satisfying all stopping criteria, the deviation is reduced to
approximately 1.5%. Extending the M&V period to match the same time
period of the conventional method (24 months) further improves accuracy
to 0.7% deviation. Furthermore, by comparing subplots (a) and (b), we
observed a minor improvement in estimation accuracy when a model is
fitted to the data to account for weather differences. The model fitting
is most beneficial for the shorter dataset (randomized and stopping when
all criteria are met, typically 24-36 weeks of data), where it reduces
deviation from 1.5% to 1.2% (around 20% improvement). This is because
the weather conditions between the first and second year for each of the
locations are relatively similar on average over this time period, and
thus the effect of adjusting for temperature as an independent variable
in the model is relatively small. We found 83% of all filtered buildings
show less error when using the randomized M&V, and by comparing the
error distribution in subplot (b), we noticed the medium deviation drops
from 4.6% to 1.2% (around 74% improvement) at the time of satisfying all
stopping criteria. Subplot (c) compares the overall timeline of the two
M&V methods. The results show that almost all buildings achieve accurate
M&V results and meet the stopping criteria within 24 weeks (for this
intervention), with the remainder meeting all stopping criteria within
36 weeks. The additional 12 weeks are usually driven by a need to span
sufficient weather conditions representative of a full year of data.
Meanwhile, the red arrow highlights the excess time required by the
conventional M&V method (24 months). Thus, the randomized M&V method
finds more accurate results in approximately a quarter of the time
required for conventional M&V. Even if it was the case that an existing
12-month baseline data was available at the start of the M&V process,
randomized M&V would still obtain a more accurate result, in less time,
than obtaining the 12-month retrofit dataset required by conventional
M&V.

@@

p_top <- df_MW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       color = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with measured weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

p_middle <- df_TW_A %>% 
  mutate(method = as.factor(method), 
         method = recode_factor(method, "conv_diff" = "Conventional", "rand_eob_diff" = "Randomized", "rand_final_diff" = "Randomized\n(24 months)")) %>% 
  ggplot(aes(x = method, y = diff, fill = method)) +
  geom_jitter(width = 0.2, alpha = 0.8, size = 0.5) +
  geom_lv(k = 4, outlier.shape = NA) +
  geom_boxplot(outlier.alpha = 0, coef = 0, fill = "#00000000", aes(color = method)) +
  geom_text(data = . %>% group_by(method) %>% summarise(median = median(diff)) %>% ungroup(), 
            aes(x = method, y = median, label = paste0(round(median, digits = 1), " %")), 
            size = 4) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 4), 
                     labels = number_format(suffix = " %")) +
  scale_fill_manual(values = ls_colors) +
  scale_color_manual(values = c("grey80", "grey80", "grey80")) + 
  labs(fill = NULL, 
       x = NULL, 
       y = "Absolute error in fractional savings", 
       subtitle = str_glue("All {A_building} buildings with TOWT model and TMY weather")) +
  coord_cartesian(ylim = c(0, 23)) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.position = "none",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

@@

ggarrange(p_accuracy, p_timeline, 
          ncol = 1, nrow = 2,
          heights = c(2.2, 1),
          labels = c("", "c)"),
          common.legend = F, 
          legend = "bottom") +
  plot_annotation(title = "Overall accuracy and time required for savings estimation by different M&V methods")

MW_summary <- df_MW_A %>% 
  pivot_wider(names_from = "method", values_from = "diff") %>% 
  group_by(name) %>% 
  summarise(improv_eob = (conv_diff - rand_eob_diff), 
            improv_final = (conv_diff - rand_final_diff)) %>% 
  ungroup() %>% 
  summarise(across(c(improv_eob, improv_final), median))

TW_summary <- df_TW_A %>% 
  pivot_wider(names_from = "method", values_from = "diff") %>% 
  group_by(name) %>% 
  summarise(improv_eob = (conv_diff - rand_eob_diff), 
            improv_final = (conv_diff - rand_final_diff), 
            improv = ifelse(improv_eob >= 0, 1, 0)) %>% 
  ungroup() %>% 
  mutate(perc = sum(improv)) %>% 
  summarise(across(c(improv_eob, improv_final, perc), median))
```

Figure \@ref(fig:meanerr) shows the biased (instead of absolute)
estimated deviation, highlighting that the uncertainty range associated
with the conventional method is significantly larger compared to even
the randomized method that only takes 36 weeks, according to subplot (c)
in Figure \@ref(fig:abstract).

```{r meanerr, fig.cap="Distribution of estimated deviation from the ground-truth savings calculated by the two M&V methods (red dashed line highlighting 0% error)", fig.height=8, fig.width=12}
# plot for the stable subset
rand_eob_S <- df_seq_fs_stable %>% 
  filter(seq == "eob") %>% 
  left_join(df_fs_stable %>% filter(scenario == "ref" & method == "true"), by = c("name", "site")) %>% 
  mutate(diff = savings - fs, 
         method = "rand_eob") %>% 
  select(name, method, diff)

@@

|                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
|-------------------|-----------------------------------|-------------------|
| 25th percentile | -5.85                                               | `r percentiles$Q1`     |
| 50th percentile | -1.25                                               | `r percentiles$Median` |
| 75th percentile | 3.86                                                | `r percentiles$Q3`     |

: Table 2. Savings accuracy comparison with a similar literature

@@

As mentioned in Section \@ref(filter-and-clean-dataset), the majority of
the selected buildings are affected by operational drift, which is the
reason why we observe baseline electricity usage changes throughout the
two-year period. To evaluate the robustness of the two M&V methods to
the effect of operational drift and non-routine events, we repeated the
M&V run on the original dataset prior to the application of the reset
control. In other words, there *is no intervention applied here*, so the
most reliable M&V method will be the one that detects the closest to 0
kW savings using the measurements. Therefore, the $Acc$ here are
calculated as:

Figure \@ref(fig:nosaving) shows that the randomized M&V method (both at
the time of satisfying all stopping criteria and at the end of a
two-year period) produced consistent results. We included the absolute
deviation ($|ACC|$) figure in the supplementary material. Similarly to
Figure \@ref(fig:abstract), stopping when meeting all stopping criteria
generally yields slightly larger uncertainty than continuing for 24
months. However, it still yields a more accurate result than the
conventional method. Interestingly, similar to
[@granderson_accuracy_2016] where they obtained a median normalized
deviation around -1.25%, our analysis on the conventional method shows
that the typical deviation is -1.5%, meaning that the typical building
is decreasing energy use slightly in this dataset in the absence of any
known intervention. Thus, if a building analyst were to apply an
intervention to this building and use conventional M&V to measure the
effect, they would typically 'detect' 1.5% energy savings even when the
intervention had no actual impact on energy performance.

@@

@@
